Hierarchical Principal Manifold Estimation

### 2023-08-16 paper/main.tex
add hierarchical pme algorithm

The embedding function which maps from the $d$-dimensional manifold to $D$-dimensional ambient space is represented as $\psi(\mathbf{y}_{ijk})$, which can be modeled as

\begin{equation}
  \mathbf{x}_{jm} = \psi_j(\mathbf{y}_{jm}, t_j) = f_p(\mathbf{y}_{jm}, t_{j}) + f_{g(j)}(\mathbf{y}_{jm}, t_{j}) + f_{i(j)}(\mathbf{y}_{jm}, t_{j}) + f_j(\mathbf{y}_{jm}) + \epsilon_{jm}. \label{eq:2}
\end{equation}

@@

\begin{algorithm}
\footnotesize
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:backfitting}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$, and $\epsilon$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$.}

@@

\begin{algorithm}
\caption{Hierarchical Principal Manifold Estimation Algorithm}\label{alg:hpme}
  \KwData{Data points $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_j}$ in $\mathbb{R}^{D}$; a positive integer $N_0 < \sum_{j=1}^{N}M_j$; $\alpha, \epsilon, \epsilon^{*} \in (0, 1)$; candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}, \left\{\gamma_p\right\}_{p=1}^{P}$; $itr \geq 1$, which is the maximum number of iterations allowed.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$, .}

>>>>

The embedding function which maps from the $d$-dimensional manifold to $D$-dimensional ambient space is represented as $\psi(\mathbf{y}_{jm})$, which can be modeled as 
\begin{equation}
  \mathbf{x}_{jm} = \psi(\mathbf{y}_{jm}, t_j) = f_p(\mathbf{y}_{jm}, t_{j}) + f_{g(j)}(\mathbf{y}_{jm}, t_{j}) + f_{i(j)}(\mathbf{y}_{jm}, t_{j}) + f_j(\mathbf{y}_{jm}) + \epsilon_{jm}. \label{eq:2}
\end{equation}
The embedding map $\psi(\mathbf{y})$ is associated with a projection index $\kappa_{\psi}(\mathbf{x}): \mathbb{R}^{D} \to \mathbb{R}^{d}$, which, informally, outputs the $d$-dimensional parameterization $\mathbf{y}$ that minimizes $\|\mathbf{x} - \psi(\mathbf{y})\|_{\mathbb{R}^{D}}$ and follows the generalized definition of projection indices given in \cite{mengPrincipalManifoldEstimation2021}.

@@

\begin{algorithm}
\footnotesize
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:backfitting}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, weights $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_j}$, group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$ and $\left\{\gamma_p\right\}_{p=1}^{P}$, $\epsilon$ and $K$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$.}

@@

\begin{algorithm}
\caption{Hierarchical Principal Manifold Estimation Algorithm}\label{alg:hpme}
  \KwData{Data points $\left\{\mathbf{v}_{jm}\right\}_{j=1, m=1}^{N, M'_j}$ in $\mathbb{R}^{D}$; group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$; a positive integer $N_0 < \min\left(\left\{M'_j\right\}_{j=1}^{N}\right)$; $\alpha, \epsilon, \epsilon^{*} \in (0, 1)$; candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}, \left\{\gamma_p\right\}_{p=1}^{P}$; $itr \geq 1$, which is the maximum number of iterations allowed.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$, .}
  \textbf{Initialization}: Apply PME with input $\left(\left\{\mathbf{v}_{1m}\right\}_{m=1}^{M'_1}, N_0, \alpha, \epsilon, \epsilon^{*}, \left\{\lambda_l\right\}_{l=1}^{L}, itr\right)$ and obtain $f^{(1)}$\;
  $\psi^{(1)} \gets f^{(1)}$\;
  \textbf{Data Reduction}: Apply HDMDE algorithm with inputs $\left(\left\{\left\{\mathbf{v}_{jm}\right\}_{m=1}^{M_j}\right\}_{j=1}^{N}, N_0, \epsilon, \alpha\right)$ and obtain $\left\{M_j\right\}_{j=1}^{N}$, $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_j}$, and $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_j}$\;
  \textbf{Parameterization}: Use the projection index of $\psi^{(1)}$, $\kappa_{\psi^{(1)}}$, to parameterize $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_j}$ by $d$-dimensional parameters $\left\{y_{jm}\right\}_{j=1, m=1}^{N, M_j}$\;
  $\mathcal{E} \gets 2 \times \epsilon^{*}$, $n \gets 1$, and $\mathcal{D}(\psi^{(1)}) \gets \sum_{j=1}^{N}\sum_{m=1}^{M_j}w_{jm}\|\mathbf{x}_{jm} - \psi^{(1)}\left(\kappa_{\psi^{(1)}}(\mathbf{x}_{jm})\right)\|_{\mathbb{R}^{D}}^2$\;
  \While{$\mathcal{E} \geq \epsilon^{*}$ and $n < itr$} {
    Compute $f_{p}^{(n + 1)}$, $\left\{f_{g}^{(n + 1)}\right\}_{g = 1}^{G}$, $\left\{f_{i}^{(n + 1)}\right\}_{i=1}^{I}$, and $\left\{f_{j}^{(n+1)}\right\}_{j=1}^{N}$ by using Algorithm \ref{alg:backfitting} with inputs $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{\kappa_{\psi^{(n)}}(\mathbf{x}_{jm})\right\}_{j=1, m=1}^{N, M_j}$, $\left\{t_j\right\}_{j=1}^{N}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{g(j)\right\}_{j=1}^{N}$, $\left\{i(j)\right\}_{j=1}^{N}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, $\epsilon$, and $K$\;
    $\left\{\psi^{(n+1)}(\mathbf{y}, t) \gets f_{p}^{(n + 1)}(\mathbf{y}, t) + f_{g(j)}^{(n+1)}(\mathbf{y}, t) + f_{i(j)}^{(n + 1)}(\mathbf{y}, t) + f_{j}^{(n+1)}(\mathbf{y})\right\}_{j=1}^{N}$\;
    $\mathcal{D}(\psi^{n+1}) \gets \sum_{j=1}^{N}\sum_{m=1}^{M_j}w_{jm}\|\mathbf{x}_{jm} - \psi^{n + 1}\left(\kappa_{\psi^{n+1}}(\mathbf{x}_{jm})\right)\|_{\mathbb{R}^{D}}^2$\;
    $\mathcal{E} \gets |\mathcal{D}(\psi^{(n+1)}) - \mathcal{D}(\psi^{(n)})| / \mathcal{D}(\psi^{(n)})$\;
    $n \gets n + 1$\;
  }
  $\hat{f}_p \gets f_{p}^{(n)}$, $\left\{\hat{f}_{g} \gets f_{g}^{(n)}\right\}_{g=1}^{G}$, $\left\{\hat{f}_{i} \gets f_{i}^{(n)}\right\}_{i=1}^{I}$, $\left\{\hat{f}_j \gets f_{j}^{(n)}\right\}_{j=1}^{N}$.
\end{algorithm}
